Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-03-30 paper/AlignedIsotropicGW.tex
DCC v2 with chi_eff bound bug fixed.
Back to red 'checkme' text after DCC posting.
Resized Figs 5 and 6

We fit hierarchical models of the three LIGO O1 observations using
these six different, zero-parameter population distributions (see
Section \ref{sec:hierarchical}).  We also fit three mixture models for
the population, where the spin magnitude distribution is fixed but the
angular distribution is a weighted sum of the isotropic and aligned
distribution.  The evidence, or marginal likelihood, for each of the
models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
distributions, the mixture models' posterior on the mixing fraction
peaks at 100\% isotropic, which explains why the zero-parameter,
pure-isotropic models are preferred over the single-parameter mixture
models for every magnitude distribution with this data set.  Not
surprisingly, given the small $\chieff$ values in the three detected
systems, the most-favoured model among those with an isotropic angular
distribution has the ``low'' magnitude distribution; the most favoured
model among those with an aligned distribution also has the ``low''
magnitude distribution.  The odds ratio between the ``low'' aligned
and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
among our suite of models.  While the data favour spin amplitude
distributions with small spin magnitudes, note that a model with all
\ac{BBH} systems having zero spin is ruled out by the GW151226
measurements, which bound $\chieff \gtrsim 0.2$ at 90\% credibility
\citep{2016PhRvL.116x1103A}.

@@

%
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{../plots/posterior_on_isotropic_fraction.png}
\caption{\textbf{Fraction of the BBH population coming from an
    isotropic distribution under a mixture model.} The dotted line
  shows the flat prior on the fraction of BBHs coming from an
  isotropic distribution, $f_i$, under the mixture model. The 3 red
  lines show the posterior on $f_i$ after O1 with our various
  assumptions regarding BH spin magnitudes.  The solid line shows the
  posterior assuming that all BHs have their spin magnitude drawn from
  the ``flat'' distribution. The dashed line assumes the ``high'' BH
  spin magnitude distribution $p(a) = 2a$. The dot-dash line assumes
  the ``low'' distribution $p(a) = 2(1-a)$.  We see that for a wide
  range of assumptions regarding BH spin magnitudes, the fraction
  coming from an isotropic distribution $f_i$ peaks at 1.}
\label{fig:mixture_fraction_posterior}
\end{figure}
%

@@

%
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{../plots/sigma_v_alpha.png}
\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin magnitude $a$
that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
\label{fig:smallspinsalpha}
\end{figure}
%

>>>>

We fit hierarchical models of the three LIGO O1 observations using
these six different, zero-parameter population distributions (see
Section \ref{sec:hierarchical}).  We also fit three mixture models for
the population, where the spin magnitude distribution is fixed but the
angular distribution is a weighted sum of the isotropic and aligned
distribution.  The evidence, or marginal likelihood, for each of the
models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
distributions, the mixture models' posterior on the mixing fraction
peaks at 100\% isotropic, which explains why the zero-parameter,
pure-isotropic models are preferred over the single-parameter mixture
models for every magnitude distribution with this data set.  Not
surprisingly, given the small $\chieff$ values in the three detected
systems, the most-favoured model among those with an isotropic angular
distribution has the ``low'' magnitude distribution; the most favoured
model among those with an aligned distribution also has the ``low''
magnitude distribution.  The odds ratio between the ``low'' aligned
and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
among our suite of models.  While the data favour spin amplitude
distributions with small spin magnitudes, note that a model with all
\ac{BBH} systems having zero spin is ruled out by the GW151226
measurements, which bound $\chieff \gtrsim 0.1$ at 90\% credibility
\citep{2016PhRvL.116x1103A}.

@@

%
\begin{figure}
\centering
\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
\caption{\textbf{Fraction of the BBH population coming from an
    isotropic distribution under a mixture model.} The dotted line
  shows the flat prior on the fraction of BBHs coming from an
  isotropic distribution, $f_i$, under the mixture model. The 3 red
  lines show the posterior on $f_i$ after O1 with our various
  assumptions regarding BH spin magnitudes.  The solid line shows the
  posterior assuming that all BHs have their spin magnitude drawn from
  the ``flat'' distribution. The dashed line assumes the ``high'' BH
  spin magnitude distribution $p(a) = 2a$. The dot-dash line assumes
  the ``low'' distribution $p(a) = 2(1-a)$.  We see that for a wide
  range of assumptions regarding BH spin magnitudes, the fraction
  coming from an isotropic distribution $f_i$ peaks at 1.}
\label{fig:mixture_fraction_posterior}
\end{figure}
%

@@

%
\begin{figure}
\centering
\includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin magnitude $a$
that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
\label{fig:smallspinsalpha}
\end{figure}
%
